Data Prep

### 2024-01-11 00.0_dataPrep.qmd
further improvements to dataPrep

-   `inpadf_touse`

@@

```{r download-inpa}
#| eval: false
# inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
# inpadf <- methods::as(inpa, "data.frame")
# inpadf_touse <- inpadf %>%
#   mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
#          year = as.numeric(lubridate::year(timestamp))) %>%
#   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
# write_feather(inpadf_touse, "data/derived/inpadf_touse.feather")

@@

-   `dat`

@@

```{r download-ornitela}
#| eval: false
# Ornitela: download data from movebank (just a subset of the times for now)
minDate <- "2020-09-01 00:00"
maxDate <- "2023-05-15 11:59"
# dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
# write_feather(ornitela, "data/derived/dat.feather")

@@

-   `dat3`

@@

inpadf_touse <- read_feather("data/derived/inpadf_touse.feather")
dat <- read_feather("data/derived/dat.feather")
# # Will there be a problem joining these?
# names(dat)[!(names(dat) %in% names(inpadf_touse))] # all names in "dat" are found in inpadf_touse
# names(inpadf_touse)[!(names(inpadf_touse) %in% names(dat))] # many are not found in "dat" that are found in inpadf_touse.
# 
# # Add the dataset names so we can keep track of where the data comes from
# inpadf_tojoin <- inpadf_touse[,names(dat)] %>%
#   mutate(dataset = "inpa")
# dat <- dat %>%
#   mutate(dataset = "ornitela")
# 
# # Join the two datasets
# joined <- bind_rows(inpadf_tojoin, dat)
# dim(joined)
# 
# ## fix trackId
# dat2 <- joined %>%
#   mutate(trackId = as.character(trackId),
#          trackId = case_when(trackId == "E03" ~ "E03w",
#                              TRUE ~ trackId))
# # number of unique individuals
# length(unique(dat2$trackId)) # 206
# 
# # Add the Nili_id
# ww <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
#   dplyr::select(Nili_id, Movebank_id) %>%
#   distinct()
# 
# # Are there any other individuals who have two different associated trackId's?
# ww %>% filter(Movebank_id %in% dat2$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
# ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's
# 
# all(dat2$trackId %in% ww$Movebank_id) # false.
# 
# # Okay, let's investigate the ones that aren't included
# dat2 %>%
#   filter(!(trackId %in% ww$Movebank_id)) %>%
#   pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces
# 
# dat2$trackId <- str_replace_all(tolower(dat2$trackId), "\\.", "")
# ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")
# 
# sort(unique(ww$mb_id))
# unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
# dat2$trackId[dat2$trackId == "y01t60w"] <- "t60white"
# dat2$trackId[dat2$trackId == "e86white"] <- "e86"
# unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
# # Investigate the remaining ones:
# dat2 %>% filter(!(trackId %in% ww$mb_id)) %>% dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>% group_by(trackId, local_identifier, ring_id, dataset) %>% summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
# # I guess for now let's just make their trackId's into Nili_id's
# 
# dat3 <- left_join(dat2, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
#                   by = c("trackId" = "mb_id"))
# nrow(dat3) == nrow(dat2) # good, same number of rows. (check that the join worked)
# length(unique(dat3$Nili_id))
# # For any that didn't get Nili_ids assigned, make it the trackId
# dat3 %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
# dat3 <- dat3 %>%
#   mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
#                              TRUE ~ Nili_id))
# unique(dat3$Nili_id)
```

@@

-   `dat3`

@@

@@

## annotate the data with periods to remove
toRemove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")

toRemove <- toRemove %>%
  dplyr::select(Nili_id,
                remove_start,
                remove_end,
                reason) %>%
  mutate(across(c(remove_start, remove_end), .fns = function(x){
    lubridate::ymd(x)
  })) %>%
  dplyr::filter(!is.na(remove_end))

@@

datAnnot_noCaptures <- datAnnot_noCaptures %>%
  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))
```

>>>>

@@

```{r download-inpa}
#| eval: false
inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
inpadf <- methods::as(inpa, "data.frame")
inpadf_touse <- inpadf %>%
  mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
         year = as.numeric(lubridate::year(timestamp))) %>%
  filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
write_feather(inpadf_touse, "data/derived/inpadf_touse.feather")

@@

@@

```{r download-ornitela}
#| eval: false
# Ornitela: download data from movebank (just a subset of the times for now)
minDate <- "2020-09-01 00:00"
maxDate <- "2023-05-15 11:59"
dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
write_feather(ornitela, "data/derived/dat.feather")

@@

-   `data/derived/dat3.feather`

```{r load-inpa-ornitela}
inpadf_touse <- read_feather("data/derived/inpadf_touse.feather")
dat <- read_feather("data/derived/dat.feather")
```

@@

# Will there be a problem joining these?
names(dat)[!(names(dat) %in% names(inpadf_touse))] # all names in "dat" are found in inpadf_touse
names(inpadf_touse)[!(names(inpadf_touse) %in% names(dat))] # many are not found in "dat" that are found in inpadf_touse.

# Add the dataset names so we can keep track of where the data comes from
inpadf_tojoin <- inpadf_touse[,names(dat)] %>%
  mutate(dataset = "inpa")
dat <- dat %>%
  mutate(dataset = "ornitela")

# Join the two datasets
joined <- bind_rows(inpadf_tojoin, dat)
dim(joined)

## fix trackId
dat2 <- joined %>%
  mutate(trackId = as.character(trackId),
         trackId = case_when(trackId == "E03" ~ "E03w",
                             TRUE ~ trackId))
# number of unique individuals
length(unique(dat2$trackId)) # 206

# Add the Nili_id
ww <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, Movebank_id) %>%
  distinct()

# Are there any other individuals who have two different associated trackId's?
ww %>% filter(Movebank_id %in% dat2$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's

all(dat2$trackId %in% ww$Movebank_id) # false.

# Okay, let's investigate the ones that aren't included
dat2 %>%
  filter(!(trackId %in% ww$Movebank_id)) %>%
  pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces

dat2$trackId <- str_replace_all(tolower(dat2$trackId), "\\.", "")
ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")

sort(unique(ww$mb_id))
unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
dat2$trackId[dat2$trackId == "y01t60w"] <- "t60white"
dat2$trackId[dat2$trackId == "e86white"] <- "e86"
unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
# Investigate the remaining ones:
dat2 %>% filter(!(trackId %in% ww$mb_id)) %>% dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>% group_by(trackId, local_identifier, ring_id, dataset) %>% summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
# I guess for now let's just make their trackId's into Nili_id's

dat3 <- left_join(dat2, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
                  by = c("trackId" = "mb_id"))
nrow(dat3) == nrow(dat2) # good, same number of rows. (check that the join worked)
length(unique(dat3$Nili_id))
# For any that didn't get Nili_ids assigned, make it the trackId
dat3 %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
dat3 <- dat3 %>%
  mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
                             TRUE ~ Nili_id))
unique(dat3$Nili_id)
write_feather(dat3, "data/derived/dat3.feather")
```

@@

-   `data/derived/dat3.feather`

@@

```{r load-hospital}
dat3 <- read_feather("data/derived/dat3.feather")
toRemove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
```

@@

## annotate the data with periods to remove
toRemove <- toRemove %>%
  dplyr::select(Nili_id,
                remove_start,
                remove_end,
                reason) %>%
  mutate(across(c(remove_start, remove_end), .fns = function(x){
    lubridate::ymd(x)
  })) %>%
  dplyr::filter(!is.na(remove_end))

@@

datAnnot_noCaptures <- datAnnot_noCaptures %>%
  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))

write_feather(datAnnot_noCaptures, "data/derived/datAnnot_noCaptures.feather")
```

## Attach age and sex information

-   `data/derived/datAnnot_noCaptures.feather`

-   `datAnnot` (`datAnnot2`)

```{r}
datAnnot_noCaptures <- read_feather("data/derived/datAnnot_noCaptures.feather")
```

```{r}
#| eval: false
# Age and sex info --------------------------------------------------------
# Attach age and sex information
as <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, birth_year, sex) %>%
  distinct()

datAnnot2 <- datAnnot_noCaptures %>%
  dplyr::select(-c("sex")) %>%
  left_join(as, by = "Nili_id")
nrow(datAnnot2) == nrow(datAnnot)
length(unique(datAnnot2$Nili_id)) # 196

datAnnot <- datAnnot2 # a hack so the rest of the code will work
```

## Clean data

-   `data/raw/CutOffRegion.kml`

-   `data/derived/datAnnotCleaned.Rda`

```{r}
#| eval: false
# Clean the data
## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
mask <- sf::st_read("data/raw/CutOffRegion.kml")
datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) # XXX using 30 days overall. If we want 30 days per season, will have to do that differently (e.g. split the seasons earlier)--but I think that that'll be covered later in the data cleaning anyway.
length(unique(datAnnotCleaned$Nili_id)) # 172 individuals
# Fix time zone so dates make sense ---------------------------------------
## Overwrite the dateOnly column from the new times
datAnnotCleaned <- datAnnotCleaned %>%
  mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
         dateOnly = lubridate::date(timestampIsrael),
         month = lubridate::month(dateOnly),
         year = lubridate::year(dateOnly))

save(datAnnotCleaned, file = "data/derived/datAnnotCleaned.Rda")
load("data/derived/datAnnotCleaned.Rda")
```

## Split into seasons

Season definitions, and code to split the data into seasons, come from Marta.

-   `datAnnotCleaned`

-   `data/derived/seasons_orig.Rda`

-   `data/derived/seasonNames_orig.Rda`

```{r}
#| eval: false
# Split into Marta's 3 seasons --------------------------------------------
s1 <- datAnnotCleaned %>%
  mutate(start_breeding = ifelse(month %in% c(1:5),
                                 paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
                                 paste(as.character(year), "-", "12", "-", "15", sep = "")),
         start_summer = ifelse(month == 12,
                               paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
                               paste(as.character(year), "-", "05", "-", "15", sep = "")),
         start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
         start_breeding = as.Date(start_breeding),
         start_summer = as.Date(start_summer),
         season = case_when(
           dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
           dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
           dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
  filter(dateOnly != "2023-05-15") # remove May 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.

s2 <- s1 %>%
  mutate(age = year - birth_year,
         sex = case_when(sex == "f/m" ~ "u",
                         TRUE ~ sex)) %>%
  mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
  group_by(Nili_id, year, month) %>%
  mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
  mutate(age_group = case_when(age == 0 ~ "juv",
                               age >= 1 & age <= 4 ~ "sub",
                               age >= 5 ~ "adult",
                               TRUE ~ NA)) %>%
  mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
                                  season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
                                  season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
  # explicitly set the factor levels in temporal order
  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding")),
         season = factor(season, levels = c("breeding", "summer", "fall")))
table(s2$seasonUnique, exclude = NULL) # XXX can use this in the report--easy to read

# remove columns that were used in the season calculation
s3 <- s2 %>%
  dplyr::select(-c(month, start_breeding, start_summer, start_fall))

# Separate the seasons -----------------------------
seasons_orig <- s3 %>%
  group_by(seasonUnique) %>%
  group_split(.keep = T)
# extract the season names in case we need a separate vector at some point.
seasonNames_orig <- map_chr(seasons_orig, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
save(seasons_orig, file = "data/derived/seasons_orig.Rda")
save(seasonNames_orig, file = "data/derived/seasonNames_orig.Rda")
load("data/derived/seasons_orig.Rda")
load("data/derived/seasonNames_orig.Rda")
```

## Remove individuals with low fix rate

Find the mode (most common) observation frequency for each individual on each day. Then exclude individuals who never have observations at 10-minute frequencies.

We keep the ones that theoretically have a 10-minute fix rate, even if in practice they often have much less frequent observations.

-   `seasons_orig`

-   `data/derived/toKeep_fixRate.Rda`

-   `seasons_mode10`

```{r}
#| eval: false
# Optionally remove individuals with a low fix rate -----------------------
Mode <- function(x) { # function to calculate mode fix rate
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Calculate daily modes and add them to the dataset
seasons <- map(seasons_orig, ~.x %>%
                 group_by(dateOnly, Nili_id) %>%
                 mutate(diff = as.numeric(difftime(lead(timestamp), timestamp, units = "mins"))) %>%
                 group_by(dateOnly, Nili_id) %>%
                 mutate(mode = Mode(round(diff))) %>%
                 ungroup())

# Exclude individuals that never have a daily mode of 10 minutes
toKeep_fixrate <- map(seasons, ~.x %>%
                        sf::st_drop_geometry() %>%
                        group_by(Nili_id) %>%
                        summarize(minmode = min(mode, na.rm = T)) %>%
                        filter(minmode <= 10) %>%
                        pull(Nili_id) %>%
                        unique())
save(toKeep_fixrate, file = "data/derived/toKeep_fixrate.Rda")
load("data/derived/toKeep_fixrate.Rda")
seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y))
```

## Save dataset for social analysis

The next step will be to remove individuals with too few points per day or too few days tracked. But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric. So, before I filter for ppd and for days tracked, going to save a copy to use for social analysis.

-   `seasons`

-   `seasons_mode10`

-   `data/derived/seasons_forSoc.Rda`

-   `data/derived/seasons_forSoc_mode10.Rda`

```{r}
# Save copies for social analysis
# 
seasons_forSoc <- seasons
seasons_forSoc_mode10 <- seasons_mode10
save(seasons_forSoc, file = "data/derived/seasons_forSoc.Rda")
save(seasons_forSoc_mode10, file = "data/derived/seasons_forSoc_mode10.Rda")
load("data/derived/seasons_forSoc.Rda")
load("data/derived/seasons_forSoc_mode10.Rda")
seasons_mode10 <- seasons_forSoc_mode10
seasons <- seasons_forSoc

# Investigate: How many individuals are retained in each of these cases? What proportion were removed?
inds_all <- map_dbl(seasons_forSoc, ~length(unique(.x$Nili_id)))
inds_mode10 <- map_dbl(seasons_forSoc_mode10, ~length(unique(.x$Nili_id)))
propRemoved <- round((inds_all-inds_mode10)/inds_all, 3) # For all seasons, getting rid of individuals that don't have 
propRemoved # yikes that's a lot of birds. Alas.
```

## Get roosts

-   `seasons`

-   `seasons_mode10`

-   `data/derived/roosts_seasons.Rda`

-   `data/derived/roosts_seasons_mode10.Rda`

```{r}
#| eval: false
# Get roosts for each season ----------------------------------------------
roosts_seasons <- purrr::map(seasons, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
roosts_seasons <- roosts_seasons %>%
  map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
save(roosts_seasons, file = "data/derived/roosts_seasons.Rda")
load("data/derived/roosts_seasons.Rda")

roosts_seasons_mode10 <- purrr::map(seasons_mode10, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
roosts_seasons_mode10 <- roosts_seasons_mode10 %>%
  map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
save(roosts_seasons_mode10, file = "data/derived/roosts_seasons_mode10.Rda")
load("data/derived/roosts_seasons_mode10.Rda")
```

## Remove nighttime points

XXX do we not also remove nighttime points for seasons_mode10?

-   `seasons`

-   `seasons`

```{r}
#| eval: false
# Remove nighttime points -------------------------------------------------
before <- map_dbl(seasons, nrow)
seasons <- map(seasons, ~{
  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
                                     lat = 31.434306, lon = 34.991889, # XXX revisit this!
                                     keep = c("sunrise", "sunset")) %>%
    dplyr::select("dateOnly" = date, sunrise, sunset)

  .x <- .x %>%
    dplyr::mutate(dateOnly = lubridate::ymd(dateOnly)) %>%
    dplyr::left_join(times, by = "dateOnly") %>%
    dplyr::mutate(daylight = ifelse(timestamp >= sunrise & timestamp <= sunset, "day", "night")) %>%
    dplyr::select(-c(sunrise, sunset)) %>%
    dplyr::filter(daylight == "day")
})
after <- map_dbl(seasons, nrow)
change <- after-before
changeDF <- data.frame(seasonName = seasonNames_orig, nBefore = before, nAfter = after) %>%
  mutate(propRowsRemoved = round((before-after)/before, 2))
changeDF
```

## Restrict to Southern individuals

Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, <https://epsg.io/32636)>

XXX why are we using `seasons_orig` here? shouldn't it be `seasons`?

-   `seasons_orig`

-   `data/derived/seasonsSF.Rda`

-   `seasons` XXX seems wrong!

```{r}
#| eval: false
# Restrict to southern individuals ----------------------------------------
# 
# Transform to SF object, so we can get centroids
seasonsSF <- map(seasons_orig, ~.x %>%
                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
                   sf::st_set_crs("WGS84") %>%
                   sf::st_transform(32636))

save(seasonsSF, file = "data/derived/seasonsSF.Rda")
load("data/derived/seasonsSF.Rda")

## Get centroids, so we can see who's "southern" for that season.
centroids <- map(seasonsSF, ~.x %>%
                   group_by(Nili_id) %>%
                   summarize(geometry = st_union(geometry)) %>%
                   st_centroid())

## Examine a histogram of centroid latitudes 
walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.

## Get southern individuals for each season, so we can filter the data
southernIndivs <- map(centroids, ~.x %>%
                        filter(st_coordinates(.)[,2] < 3550000) %>%
                        pull(Nili_id))

## Remove individuals not in the south
before <- seasons
seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y))
after <- seasons

beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
df <- data.frame(season = seasonNames_orig, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
df
rm(centroids)
rm(seasonsSF)
```

## Remove individuals with too few points per day

-   `seasons`

-   `seasons`

```{r}
#| eval: false
# Include only individuals with enough points per day ------------------------
# Investigate battery percentage vs points per day ------------------------
# Is there some way we can relate battery charge percentage to number of points per day?
test <- seasons[[5]]
names(test)

ppd <- map_dfr(seasons, ~.x %>%
                 st_drop_geometry() %>%
                 group_by(seasonUnique, season, Nili_id, dateOnly) %>%
                 summarize(n = n(),
                           minBatt = min(battery_charge_percent, na.rm = T),
                           meanBatt = mean(battery_charge_percent, na.rm = T),
                           medBatt = mean(battery_charge_percent, na.rm = T)))

excl_batt <- 50
excl_pts <- 10
## min battery charge
ppd %>%
  filter(n < 100) %>% # only interested in the lower numbers
  mutate(remove = case_when(minBatt <= excl_batt & n <= excl_pts ~ T,
                            TRUE ~ F)) %>%
  ggplot(aes(x = minBatt, y = n))+
  geom_point(aes(col = remove))+
  facet_wrap(~seasonUnique) +
  scale_color_manual(values = c("black", "red")) +
  theme_classic()

# # cumulative histogram to look at the results of removing days based only on ppd
# ppd %>%
#   filter(n < 100) %>%
#   ggplot(aes(x = n)) +
#   stat_ecdf(aes(y = after_stat(y)*100))+
#   ylab("Percentage of vulture-days included")+
#   xlab("Points per day")+
#   theme_classic()+
#   facet_wrap(~seasonUnique)
# 
# # cumulative histogram to look at results of removing days based on ppd after battery is taken into account (min pct < 50)
# ppd %>%
#   filter(n < 100, minBatt > 50) %>%
#   ggplot(aes(x = n)) +
#   stat_ecdf(aes(y = after_stat(y)*100))+
#   ylab("Percentage of vulture-days included")+
#   xlab("Points per day")+
#   theme_classic()+
#   facet_wrap(~seasonUnique)
# 
# # same thing, but with batery set at 75
# # cumulative histogram to look at results of removing days based on ppd after battery is taken into account (min pct < 50)
# ppd %>%
#   filter(n < 100, minBatt > 75) %>%
#   ggplot(aes(x = n)) +
#   stat_ecdf(aes(y = after_stat(y)*100))+
#   ylab("Percentage of vulture-days included")+
#   xlab("Points per day")+
#   theme_classic()+
#   facet_wrap(~seasonUnique)
# 
# # What if instead we decide we want a certain number of ppd--battery cuttoff may be sharper. Say, 30 ppd: is there a discernible battery cutoff that distinguishes "real" from "battery artifact" days?
# ppd %>%
#   filter(n < 30) %>%
#   ggplot(aes(x = minBatt))+
#   stat_ecdf(aes(y = after_stat(y)*100))+
#   ylim(c(0, 100))+
#   facet_wrap(~seasonUnique)+
#   ylab("Percentage of vulture-days included")+
#   xlab("Minimum battery percentage") # we do seem to see crooks here around 75%!
# 
# # Same thing but with 10 points per day? Predict much sharper cutoffs.
# ppd %>%
#   filter(n < 10) %>%
#   ggplot(aes(x = minBatt))+
#   stat_ecdf(aes(y = ..y..*100))+
#   ylim(c(0, 100))+
#   facet_wrap(~seasonUnique)+
#   ylab("Percentage of vulture-days included")+
#   xlab("Minimum battery percentage") # crooks around 50% battery minimum for 10ppd. This seems like a pretty good way to go!

# Let's throw out anything with < 10ppd unless the battery has a min charge > 50%.

# Orr suggests to just restrict to vulture*days with >30 points per day. Let's see what that would look like.
thresh <- 30
# ppd %>%
#   mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
#   ggplot(aes(x = n))+
#   geom_histogram(aes(fill = Status))+
#   facet_wrap(~seasonUnique, scales ="free")+
#   geom_vline(aes(xintercept = thresh), col = "red")+
#   scale_fill_manual(values = c("darkgray", "red"))+
#   theme_classic()+
#   ggtitle(paste("n removed =", thresh, sep = " "))+
#   ylab("# vulture*days")+
#   xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
# 
# ppd <- ppd %>%
#   ungroup() %>%
#   mutate(removed = ifelse(n < thresh, T, F))
# 
# # what proportion of vulture*days are removed overall?
# (propRemoved <- ppd %>%
#     group_by(seasonUnique, season) %>%
#     summarize(nTot = n(),
#               nRemoved = sum(removed),
#               propRemoved = nRemoved/nTot))
# propRemoved %>%
#   ggplot(aes(x = seasonUnique, y = propRemoved, col = season))+
#   geom_point(size = 10)+
#   theme_classic()+
#   scale_color_manual(values = c("blue", "red", "darkorange"))
# 
# # I'm a bit concerned about this because it removes more data for the breeding season than for the non-breeding seasons (likely because less sunlight during those seasons)
# 
# # what about by individual over each season?
# ppd %>%
#   group_by(seasonUnique, season, Nili_id) %>%
#   summarize(propRemoved = sum(removed)/n()) %>%
#   ggplot(aes(x = seasonUnique, y = propRemoved))+
#   geom_boxplot(aes(group = seasonUnique, fill = season))+
#   geom_jitter(width = 0.2, alpha = 0.3, size = 2)+
#   scale_fill_manual(name = "Season", values = c("blue", "red", "darkorange"))+
#   theme_classic()+ # tells the same story
#   ylab("Proportion of days removed")+
#   xlab("")

# Based on the battery charge analysis above, going to remove individuals that have 10 or fewer points per day, but only if the minimum battery charge is less than 50%.

## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50%
seasons <- map(seasons, ~.x %>%
                 group_by(Nili_id, dateOnly) %>%
                 filter(n() >= 10 | (n() < 10 & min(battery_charge_percent, na.rm = T) > 50)) %>%
                 ungroup())
```

## Remove individuals with too few days tracked

-   `seasons`

-   `seasons`

```{r}
#| eval: false
# Include only individuals with enough days tracked -----------------------
## Must be tracked for at least 30 days per season
durs <- map_dbl(seasons, ~length(unique(.x$dateOnly)))

# walk2(seasons, durs, ~.x %>%
#         st_drop_geometry() %>%
#         group_by(Nili_id) %>%
#         summarize(nDaysTracked = length(unique(dateOnly)),
#                   propDaysTracked = nDaysTracked/.y) %>%
#         pull(propDaysTracked) %>%
#         hist())

indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>%
                       st_drop_geometry() %>%
                       group_by(Nili_id) %>%
                       summarize(nDaysTracked = length(unique(dateOnly)),
                                 propDaysTracked = nDaysTracked/.y) %>%
                       filter(nDaysTracked >= 30) %>%
                       pull(Nili_id))

## remove those individuals not tracked for enough days
before <- seasons
after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y))

beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
seasons <- after
```

## Calculate flight altitude

-   `seasons`

-   `data/derived/elevs_z10_seasons.Rda`

-   `data/derived/groundElev_z10.Rda`

-   `seasons`

```{r}
#| eval: false
# Get elevation rasters ---------------------------------------------------
# Remove the first element of `seasons` because it doesn't have any data
seasons <- seasons[-1] # removing summer 2020
# # XXXXXXX start here
#seasons <- map(seasons, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84"))
#elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x , z = 10))
#save(elevs_z10_seasons, file = "data/derived/elevs_z10_seasons.Rda")
load("data/derived/elevs_z10_seasons.Rda")

# groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y)) # have to convert the crs because the raster from elevatr is in WGS84.
# save(groundElev_z10, file = "data/derived/groundElev_z10.Rda")
load("data/derived/groundElev_z10.Rda")

## use elevation rasters to calculate height above ground level
before <- seasons
seasons <- map2(.x = seasons, .y = groundElev_z10, 
                ~.x %>% mutate(groundElev = .y) %>%
                  mutate(height_above_ground = height_above_msl-groundElev,
                         height_above_ground = case_when(height_above_ground < 0 ~ 0,
                                                         TRUE ~ height_above_ground)))
after <- seasons

# There should be no change in the number of rows
all(map2_lgl(before, after, ~nrow(.x) == nrow(.y)))
```

## Restrict to mode10 individuals

XXX what is happening here

-   `data/derived/toKeep_fixrate.Rda`

-   `seasons`

-   `data/derived/seasons.Rda`

-   `data/derived/seasons_mode10.Rda`

```{r}
#| eval: false
# Now restrict to the mode10 individuals ----------------------------------
load("data/derived/toKeep_fixrate.Rda")
toKeep_fixrate <- toKeep_fixrate[-1]
seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y))

save(seasons, file = "data/derived/seasons.Rda")
save(seasons_mode10, file = "data/derived/seasons_mode10.Rda")
load("data/derived/seasons.Rda")
load("data/derived/seasons_mode10.Rda")
```

## Remove high-frequency fixes

Downsampling data to 10-minute intervals exclusively

-   `seasons`

-   `seasons_mode10`

-   `data/derived/seasons_10min.Rda`

-   `data/derived/seasons_mode10_10min.Rda`

XXX this is confusing. Need to rename after concepts, not minutes.

```{r}
#| eval: false
# Downsample the data and save the downsampled data
subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){
  sub <- df %>%
    arrange(idCol, timestampCol) %>%
    group_by(.data[[idCol]]) %>%
    mutate(tc = cut(.data[[timestampCol]], 
                    breaks = paste(as.character(mins), "min", sep = " "))) %>%
    group_by(.data[[idCol]], 
             "d" = lubridate::date(.data[[timestampCol]]), tc) %>%
    slice(1) %>%
    ungroup() %>%
    dplyr::select(-c("d", "tc"))
  return(sub)
}

# create and save downsampled datasets for later use
seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling.
seasons_mode10_10min <- map(seasons_mode10, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)

map_dbl(seasons, nrow)
map_dbl(seasons_10min, nrow)
map_dbl(seasons_mode10, nrow)

seasons_mode10_10min <- map(seasons_mode10_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly))))
seasons_10min <- map(seasons_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly))))
